---
title: "Tennis Surface Report"
author: "Siddharth Gowda"
date: "7/26/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Desktop/Moneyball")
library(tidyverse)
library(plyr)
library(janitor)
library(devtools)
library(modelr)
library(Hmisc)
```

# Introduction

It's well known that different tennis court surfaces have unique 
properties, and those properties affect certain aspects of the game 
differently. Thus, for this project we analyzed singles ATP match data from 2019,
and performed analysis, comparing data from each surface type.

## Background

Below is a data frame describing the general differences about
each surface.


```{r}
surface_characteristics = 
  read.csv("~/Desktop/Moneyball/Data/surfaces_elements.csv")

head(surface_characteristics)

```

What we mean by speed: Speed describes the pace of the game. Generally on clay
courts points last longer and baseline play is used more often than compared to
other courts. This is followed by hard court and then grass, where numerous
players serve and volley, leading to points ending relatively quickly.

What we mean by spin: The spin variable describes the type of spin that is
generally accosted with the surface type. In reality all shots of these spin
types are used across all surfaces.


What we mean by strategy: Strategy measures the common methodology players
incorporate on each surface. Obviously, this isn't an exact science because
players will want to play to their strengths and their opponents weakness.
Nevertheless, generally players are more tactical on clay surfaces, increasing
the variety in there shot selection. Hard court is a mixture of strategy types,
but players do tend to brute-force and out-muscle opponents with powerful shots; 
grass strategy on the other hand is typically serve and volley.


What we mean by risk-factor: Risk-factor measures amount of risk 
(or margin of error) players allow on there strokes. For instance, 
when a player is trying to hit groundstroke down the line, if they aim exactly 
for the singles-ally line that would indicate a lower margin of error 
(or more risk), whereas if the player aimed a couple feet inside of the line 
that would be a shot of a higher margin of error (or less risk). Typically, 
players take riskier shots on grass and hard courts when compared with clay.

Extra Notes: In reality every surface type uses different types of tactics, spin,
risk, and speed (it mostly depends on the player); however, we are describing
each surface in the simplest manner. Also, we are only going to be analyzing
the major surface types (clay, grass, and hard), not specifics (i.e. green clay
vs red clay).


## Loading Data

For this analysis we used 2019 ATP singles match data via Jeff Sackmann on Github.


```{r}
atp_matchies_2019_OG = read.csv("./Data/atp_matches_2019.csv")


head(atp_matchies_2019_OG)
```

# Data Cleaning

```{r}
atp_data_rough = clean_names(atp_matchies_2019_OG) %>%
  mutate(total_games = l_sv_gms + w_sv_gms,
         w_bp_att = l_bp_faced, l_bp_att = w_bp_faced,
         w_total_pts = w_svpt + l_svpt,
         l_total_pts = w_svpt + l_svpt) %>% 
  select(tourney_name, surface, tourney_level, 
         winner_name, winner_hand, winner_ht, winner_age, c(19:21, 23:46, 48), 
         winner_rank_points, loser_rank_points, winner_seed, loser_seed, 
         total_games, w_bp_att, l_bp_att,  w_total_pts, l_total_pts)

head(atp_data_rough)

```

There were numerous unnecessary variables (i.e. tourney_ID) that we decided 
to remove. Moreover, the data was not distributed by player match stats,
instead there were variables regarding the winner and the loser. Therefore, in
preparation for separating the data by player we had to rename and add column 
types like w_bp_att (winner break points attempted) to l_bp_faced (loser break
points faced).

## Creating New Data frames

```{r}

win_atp_data = atp_data_rough[, c(1:7, 12:24, 34, 36, 38, 40, 41, 43)]

#head(win_atp_data)

loser_atp_data = atp_data_rough[, c(1:3, 8:15, 25:33, 35, 37, 39, 40, 42, 44)]

#head(loser_atp_data)

```

```{r}
win_atp_data = win_atp_data %>% rename(c("winner_name" = "name", 
                                         "winner_hand" = "hand", 
                                         "winner_ht" = "height", 
                                         "winner_age" = "age",
                                         "w_ace" = "ace", "w_df" = "df", 
                                         "w_svpt" = "svpt",
                                         "w_1st_in" = "first_in", 
                                         "w_1st_won" = "first_won", 
                                         "w_2nd_won" = "second_won", 
                                         "w_sv_gms" = "sv_gms",
                                         "w_bp_saved" = "bp_saved", 
                                         "winner_rank" = "rank",
                                         "w_bp_faced" = "bp_faced", 
                                         "winner_rank_points" = "rank_pts", 
                                         "winner_seed" = "tourney_seed",
                                         "w_bp_att" = "bp_att", 
                                         "w_total_pts" = "total_pts")) %>% 
  mutate(winner = 1)

loser_atp_data = loser_atp_data %>% rename(c("loser_name" = "name", 
                                             "loser_hand" = "hand", 
                                         "loser_ht" = "height", 
                                         "loser_age" = "age",
                                         "l_ace" = "ace", "l_df" = "df", 
                                         "l_svpt" = "svpt",
                                         "l_1st_in" = "first_in", 
                                         "l_1st_won" = "first_won", 
                                         "l_2nd_won" = "second_won", 
                                         "l_sv_gms" = "sv_gms",
                                         "l_bp_saved" = "bp_saved", 
                                         "loser_rank" = "rank", 
                                         "l_bp_faced" = "bp_faced", 
                                         "loser_rank_points" = "rank_pts", 
                                         "loser_seed" = "tourney_seed",
                                         "l_bp_att" = "bp_att", 
                                         "l_total_pts" = "total_pts")) %>% 
  mutate(winner = 0)

head(win_atp_data)
head(loser_atp_data)
```

Here we had to rename all of the columns to the same name in order to
bind the two data frames together.

```{r}
atp_data = rbind.data.frame(win_atp_data, loser_atp_data)
```

Here we bound the two data frames together.

## Binarization and other cleaning 

```{r}
atp_data = atp_data %>% 
  mutate(hand = if_else(hand == "R", 1, 0)) %>% 
  arrange(name) %>%
  filter(rank <= 100)
atp_data_no_ret = atp_data[!grepl('RET|W/O', atp_data$score),]

```

We filter the data to only include players that were at one time in the top 100. 
Moreover, we took out matches that ended in retirement because they 
were not representative a typical ATP match. This was because we wanted a sample 
of individuals who were similar, but  relatively diverse at the same time. 
We binarized the data (i.e. hand dominance and win) just in case 
we wanted to perform tests or models that needed data in that format.

## Grouping Data by indivudal

```{r}
atp_data_player = atp_data_no_ret %>% 
  group_by(name) %>% 
  dplyr::summarise(mean_win_rate = mean(winner, na.rm = TRUE),
                   mean_fs_win_rate = sum(first_won, na.rm = TRUE)/sum(first_in, na.rm = TRUE), 
                   mean_sec_srv_rate = mean((svpt - first_in)/svpt, na.rm = TRUE),
                   mean_sec_win_rate = sum(second_won, na.rm = TRUE)/sum(svpt - first_in, na.rm = TRUE),
                   mean_ace_rate = sum(ace, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   mean_pts_min = sum(total_pts, na.rm = TRUE)/sum(minutes, na.rm = TRUE),
                   mean_bp_faced_rate = sum(bp_faced, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   mean_bp_att_rate = sum(bp_att, na.rm = TRUE)/sum(total_pts - svpt, na.rm = TRUE), 
                   mean_bp_save_rate = sum(bp_saved, na.rm = TRUE)/sum(bp_faced, na.rm = TRUE),
                   mean_age = mean(age, na.rm = TRUE),
                   mean_df_rate = sum(df, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   hand = mean(hand),
                   mean_rank = mean(rank, na.rm = T),
                   mean_svpt = mean(svpt, na.rm = T)) %>%
  ungroup()

#write.csv(atp_data_player, "./Data/atp_data_player.csv")

atp_data_hard = atp_data_no_ret %>% filter(surface == "Hard") %>% group_by(name) %>%
  dplyr::summarise(mean_win_rate = mean(winner, na.rm = TRUE),
                   mean_fs_win_rate = sum(first_won, na.rm = TRUE)/sum(first_in, na.rm = TRUE), 
                   mean_sec_srv_rate = mean((svpt - first_in)/svpt, na.rm = TRUE),
                   mean_sec_win_rate = sum(second_won, na.rm = TRUE)/sum(svpt - first_in-df, na.rm = TRUE),
                   mean_sec_in_rate = sum(svpt - first_in-df, na.rm = TRUE)/sum(svpt,na.rm = TRUE),
                   mean_ace_rate = sum(ace, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   mean_pts_min = sum(total_pts, na.rm = TRUE)/sum(minutes, na.rm = TRUE),
                   mean_bp_faced_rate = sum(bp_faced, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   mean_bp_att_rate = sum(bp_att, na.rm = TRUE)/sum(total_pts - svpt, na.rm = TRUE), 
                   mean_bp_save_rate = sum(bp_saved, na.rm = TRUE)/sum(bp_faced, na.rm = TRUE),
                   mean_age = mean(age, na.rm = TRUE),
                   mean_df_rate = sum(df, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   hand = mean(hand),
                   mean_rank = mean(rank, na.rm = T)) %>%
  ungroup()%>% 
    mutate(surface = "Hard")

#write.csv(atp_data_hard, "./Data/atp_data_hard.csv")


  atp_data_clay = atp_data_no_ret %>% filter(surface == "Clay") %>%
    group_by(name) %>% 
    dplyr::summarise(mean_win_rate = mean(winner, na.rm = TRUE),
                   mean_fs_win_rate = sum(first_won, na.rm = TRUE)/sum(first_in, na.rm = TRUE), 
                   mean_sec_srv_rate = mean((svpt - first_in)/svpt, na.rm = TRUE),
                   mean_sec_win_rate = sum(second_won, na.rm = TRUE)/sum(svpt - first_in-df, na.rm = TRUE),
                   mean_sec_in_rate = sum(svpt - first_in-df, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   mean_ace_rate = sum(ace, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   mean_pts_min = sum(total_pts, na.rm = TRUE)/sum(minutes, na.rm = TRUE),
                   mean_bp_faced_rate = sum(bp_faced, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   mean_bp_att_rate = sum(bp_att, na.rm = TRUE)/sum(total_pts - svpt, na.rm = TRUE), 
                   mean_bp_save_rate = sum(bp_saved, na.rm = TRUE)/sum(bp_faced, na.rm = TRUE),
                   mean_age = mean(age, na.rm = TRUE),
                   mean_df_rate = sum(df, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   hand = mean(hand),
                   mean_rank = mean(rank, na.rm = T)) %>%
  ungroup()%>% 
    mutate(surface = "Clay")

#write.csv(atp_data_clay, "./Data/atp_data_clay.csv")

  atp_data_grass = atp_data_no_ret %>% filter(surface == "Grass")%>% 
    group_by(name) %>% 
    dplyr::summarise(mean_win_rate = mean(winner, na.rm = TRUE),
                   mean_fs_win_rate = sum(first_won, na.rm = TRUE)/sum(first_in, na.rm = TRUE), 
                   mean_sec_srv_rate = mean((svpt - first_in)/svpt, na.rm = TRUE),
                   mean_sec_win_rate = sum(second_won, na.rm = TRUE)/sum(svpt - first_in-df, na.rm = TRUE),
                   mean_sec_in_rate = sum(svpt - first_in-df, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   mean_ace_rate = sum(ace, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   mean_pts_min = sum(total_pts, na.rm = TRUE)/sum(minutes, na.rm = TRUE),
                   mean_bp_faced_rate = sum(bp_faced, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   mean_bp_att_rate = sum(bp_att, na.rm = TRUE)/sum(total_pts - svpt, na.rm = TRUE), 
                   mean_bp_save_rate = sum(bp_saved, na.rm = TRUE)/sum(bp_faced, na.rm = TRUE),
                   mean_age = mean(age, na.rm = TRUE),
                   mean_df_rate = sum(df, na.rm = TRUE)/sum(svpt, na.rm = TRUE),
                   hand = mean(hand),
                   mean_rank = mean(rank, na.rm = T)) %>%
  ungroup() %>% 
    mutate(surface = "Grass")
  
#write.csv(atp_data_grass, "./Data/atp_data_grass.csv")
```

We decided to group the data by players and create rate and mean stats. This was
because if we did not use means, better players, who generally go farther in
tournaments and play more matches would be weighted higher in our analysis
than lower ranked players. We also created rate stats to control for the fact 
that some players due to their play style would play longer matches than others.



# Performing the Analysis

In this section we will perform both EDA analysis and create models.

## Re-loading Data

```{r}

atp_data_hard = read.csv("~/Desktop/Moneyball/Data/atp_data_hard.csv") %>% 
  mutate(surface == hard) %>% 
  select(!X)
atp_data_grass = read.csv("~/Desktop/Moneyball/Data/atp_data_grass.csv") %>% 
  mutate(surface == grass) %>% 
  select(!X)

atp_data_clay = read.csv("~/Desktop/Moneyball/Data/atp_data_clay.csv")

atp_data_clay = atp_data_clay %>% 
  mutate(surface == clay) %>% 
  select(!X)

atp_data_og = read.csv("~/Desktop/Moneyball/Data/atp_data_2019.csv") 

atp_data_player = read.csv("~/Desktop/Moneyball/Data/atp_data_player.csv")

atp_data_player = atp_data_player %>% 
  select(-X)


atp_data_mlm = rbind.data.frame(atp_data_hard, atp_data_grass, atp_data_clay)
```

We created the atp_data_mlm data frame in order to distinguish the differences
between each surface type by creating an aggregate model.

## Box Plots

```{r}

ggplot(atp_data_mlm) +
  geom_boxplot(aes(mean_fs_win_rate, color = surface)) +
  labs(x = "Mean First Serve Win Rate") +
  coord_flip()

ggplot(atp_data_mlm) +
  geom_boxplot(aes(mean_bp_att_rate, color = surface)) +
  labs(x = "Mean Break Point Attempt Rate") +
  coord_flip()

ggplot(atp_data_mlm) +
  geom_boxplot(aes(mean_bp_faced_rate, color = surface)) +
  labs(x = "Mean Break Point Faced Rate") +
  coord_flip()

ggplot(atp_data_mlm) +
  geom_boxplot(aes(mean_sec_win_rate, color = surface)) +
  labs(x = "Mean Second Serve Win Rate") +
  coord_flip()

ggplot(atp_data_mlm) +
  geom_boxplot(aes(mean_df_rate, color = surface)) +
  labs(x = "Mean Double Fault Rate") +
  coord_flip()


ggplot(atp_data_mlm) +
  geom_boxplot(aes(mean_sec_srv_rate, color = surface)) +
  labs(x = "Mean Second Serve Rate") +
  coord_flip()

ggplot(atp_data_mlm) +
  geom_boxplot(aes(mean_pts_min, color = surface)) +
  labs(x = "Mean Points Per Minutes") +
  coord_flip()

```
Looking at these scatter plots there are defiantly some variables that differ
on each surface type and others that stay the safe.

Some of our prior theories in the background section are supported by these 
plots. Notably, looking at points per minute statistic (ppm), generally clay 
has the fewest ppm followed by hard and grass, which aligns with our notion 
that clay is the slowest surface, then hard, and then grass.

## Creating a Model & Our Ideology

We decided to create a multivariable regression because there are numerous
variables that impact a person's performance on the tennis court. Then
we would look at the summary statistics and other graphs in order to
describe the difference in each surface type.

## Variables Chosen

1. First Serve Win Rate: The proportion of points won when the first serve
was in.

2. Break Point Attempt Rate: The amount of break points produced divided by the
total amount of points played as the returner.

3. Break Point Faced Rate: The amount of break points faced divided by the
total amount of points played as the server.

4. Second Serve Rate: The proportion of second serves attempted divided by the
number of serves attempted.

What is a break point?

A break point occurs when the returner has to win one more point to win a game
on the opponent's serve (also known as a service game). 

Examples: 0-40, 30-40, or ad-out

We chose these variables because it is known that the serve is the most
important shot in tennis (since it is the only one that is fully controllable),
and because in the ATP specifically players are expected to maintain service 
games. Thus, break points are an important factor in getting a leg-up on the 
opponent, which typically leads to more winning.


## Scatter Plots of Variables Chosen

```{r}
ggplot(atp_data_mlm) +
  geom_point(aes(mean_fs_win_rate, mean_win_rate, color = surface)) +
  labs(x = "Mean First Serve Win Rate", y = "Mean Win Percentage")

ggplot(atp_data_mlm) +
  geom_point(aes(mean_bp_att_rate, mean_win_rate, color = surface)) +
  labs(x = "Mean Break Point Attempt Rate", y = "Mean Win Percentage")

ggplot(atp_data_mlm) +
  geom_point(aes(mean_bp_faced_rate, mean_win_rate, color = surface)) +
  labs(x = "Mean Break Point Faced Rate", y = "Mean Win Percentage")

ggplot(atp_data_mlm) +
  geom_point(aes(mean_sec_srv_rate, mean_win_rate, color = surface)) +
  labs(x = "Mean Second Serve Rate", y = "Mean Win Percentage")


```
Above are the scatter plots of all of variable in our multivariable regression.
There are certainly some differences regarding each surface type, but the
scatter plot is pretty cultured so we decided to create linear models for each
variable based on surface type.

## Linear Models

Blacks mean not grouped by surface type (aggregate) and is meant to show
the problems in a one-size fits all model.

```{r}
#First Serve Win Rate with Winning Rate

ggplot(atp_data_hard) +
  geom_abline(intercept = -1.0051, slope = 2.0293, col = "Black", size = 1) +
  geom_abline(intercept = -0.6526, slope = 1.5788, col = "red", size = 1) +
  geom_abline(intercept = -1.5780, slope = 2.7394, col = "Green", size = 1) +
  geom_abline(intercept = -0.9492, slope = 1.9169, col = "blue", size = 1) +
  xlim(0.5, 1) +
  ylim(0,1) +
  labs(x = "Mean First Serve Win Rate", y = "Mean Win Percentage")

#summary(lm(mean_win_rate~mean_fs_win_rate, data = atp_data_player))
#summary(lm(mean_win_rate~mean_fs_win_rate, data = atp_data_clay))
#summary(lm(mean_win_rate~mean_fs_win_rate, data = atp_data_grass))
#summary(lm(mean_win_rate~mean_fs_win_rate, data = atp_data_hard))

```
As seen above, there is generally a positive association between first serve win
rate and winning percentage. This makes sense because someone would expect that
if a player wins more points on their first serve they will improve their 
chances at winning the match. 

The slope for grass courts is the steepest slope (2.74) and this makes 
intuitive sense, because on grass courts the game is faster and players take 
more risk. Thus, having an effective first serve that puts the server in 
an advantageous position to win the next point is crucial. Clay and hard courts 
are a bit closer together, with slopes of 1.58 and 1.92 respectively, because
the style of play is slower than that of grass. Also, the fact that clay has the
shallowest slope should be expected since due to the nature of the surface, 
points are longer and therefore, the advantages of the serve are lessened.


```{r}
#Break Point Attempt Rate with Winning Rate

ggplot(atp_data_mlm) +
  geom_abline(intercept = 0.05034, slope = 5.10571, col = "Black", size = 1) +
  geom_abline(intercept = 0.04816, slope = 4.36168, col = "red", size = 1) +
  geom_abline(intercept = 0.03427, slope = 5.58964, col = "Green", size = 1) +
  geom_abline(intercept = 0.00519, slope = 5.88986, col = "blue", size = 1) +
  xlim(0,0.25) +
  ylim(0,1) +
  labs(x = "Mean Break Point Attempt Rate", y = "Mean Win Percentage")

#summary(lm(mean_win_rate~mean_bp_att_rate, data = atp_data_player))
#summary(lm(mean_win_rate~mean_bp_att_rate, data = atp_data_clay))
#summary(lm(mean_win_rate~mean_bp_att_rate, data = atp_data_grass))
#summary(lm(mean_win_rate~mean_bp_att_rate, data = atp_data_hard))

```
The linear models show a positive correlation between break points attempted
and winning percentage. This makes sense because it is difficult to break serve
on the ATP tour. Thus, the more chances to break an opponents the serve the more
likely it is for someone to win a match.

Along with that, the fact that clay has the shallowest slope (4.36) makes 
sense because due to the nature of the surface, points are longer on clay and 
require more shots. Thus, the ability for a player to gain free points off the
serve or set themselves up to win a point in the next two or three shots is
relatively weakened on the clay court. Therefore, it will be easier to generate
break points, but that also works against the player when they are serving.
Similarly, it makes sense that the hard court and grass courts have a steeper 
slope because it is generally easier to generate aces, missed returners, or win 
a point in the next 1 to 3 shoots (i.e. the serve and volley tactic on 
grass courts).


```{r}
#Break Point Faced Rate with Winning Rate

ggplot(atp_data_mlm) +
  #geom_point(aes(mean_bp_faced_rate, mean_win_rate, color = surface)) +
  geom_abline(intercept = 0.7947, slope = -4.1248, col = "Black", size = 1) +
  geom_abline(intercept = 0.83126, slope = -4.16520 , col = "red", size = 1) +
  geom_abline(intercept = 0.92894, slope = -6.58008, col = "Green", size = 1) +
  geom_abline(intercept = 0.75574, slope = -3.99290 , col = "blue", size = 1) +
  xlim(0,0.25) +
  ylim(0,1) +
  labs(x = "Mean Break Point Faced Rate", y = "Mean Win Percentage")

#summary(lm(mean_win_rate~mean_bp_faced_rate, data = atp_data_player))
#summary(lm(mean_win_rate~mean_bp_faced_rate, data = atp_data_clay))
#summary(lm(mean_win_rate~mean_bp_faced_rate, data = atp_data_grass))
#summary(lm(mean_win_rate~mean_bp_faced_rate, data = atp_data_hard))

```
There is negative correlation between break points faced and win percentage.
This should be the case because as previously mentioned ATP players are
expected to win their service games. Hence, if a players faces many break points
they are more likely to get broken. That notion combined with the fact that it is
generally difficult to break ATP players' service games means, if a player 
faces many break points they are more likely to lose the match.

The fact that grass court slope is the steepest (-6.58) makes sense, because
generally it is easier to maintain serve on grass, because of the way the ball
skids on a grass court. Thus, if someone has a nasty slice serve or a good serve
in general, they will generally win the service game. Therefore, if someone
gets broken it is generally even more difficult to break back than on other 
surfaces.

As for clay, it makes sense that its slope is less than grass (-4.17) because
as mentioned before, more break points are generated on clay. It is actually
interesting that hard court has the shallowest slope (-4), and this is something
that we would further study if given more time. Also, perhaps the difference 
between clay and hard court is not statistically significant.


```{r}
#Second Serve Rate Rate with Winning Rate

ggplot(atp_data_mlm) +
  #geom_point(aes(mean_sec_srv_rate, mean_win_rate, color = surface)) +
  geom_abline(intercept = 0.8832, slope = -1.1479, col = "Black", size = 1) +
  geom_abline(intercept = 0.9186, slope = -1.2676 , col = "red", size = 1) +
  geom_abline(intercept = 0.6182, slope = -0.5090, col = "Green", size = 1) +
  geom_abline(intercept = 0.6732, slope = -0.6299 , col = "blue", size = 1) +
  xlim(0,1) +
  ylim(0,1) +
  labs(x = "Mean Second Serve Rate", y = "Mean Win Percentage")

#summary(lm(mean_win_rate~mean_sec_srv_rate, data = atp_data_player))
#summary(lm(mean_win_rate~mean_sec_srv_rate, data = atp_data_clay))
#summary(lm(mean_win_rate~mean_sec_srv_rate, data = atp_data_grass))
#summary(lm(mean_win_rate~mean_sec_srv_rate, data = atp_data_hard))

```
As seen above, there is a negative association between second serve rate and
win percentage. This makes sense because on second serves players have
to be less aggressive and more risk-adverse, resulting in serves that are easier
for the opponents to attack.

This idea is exacerbated on clay because players are more risk-adverse on clay
court because they know that generally they won't be winning points on the serve
or set themselves up for two or three shot points. As for hard and grass,
the slopes are lower because serves are not as easily attacked as they are on
clay.

# Results

```{r}
multi_lm_atp_hard = lm(mean_win_rate ~ mean_fs_win_rate + mean_bp_faced_rate + mean_bp_att_rate + mean_sec_srv_rate, data = atp_data_hard)

summary(multi_lm_atp_hard)


multi_lm_atp_grass = lm(mean_win_rate ~ mean_fs_win_rate + mean_bp_faced_rate + mean_bp_att_rate + mean_sec_srv_rate, data = atp_data_grass)

summary(multi_lm_atp_grass)

multi_lm_atp_clay = lm(mean_win_rate ~ mean_fs_win_rate + mean_bp_faced_rate + mean_bp_att_rate + mean_sec_srv_rate, data = atp_data_clay)

summary(multi_lm_atp_clay)
```

# Evaluating the Results

All p-values were formed against the null of a coefficient of zero.

## Hard Court

The most statistically significant variables for hard court were
1. Break Point Attempt rate
2. Break Point Faced Rate
3. First Serve Win Rate

## Grass Court

The most statistically significant variables for grass court were
1. Break Point Attempt Rate
2. Break Point Faced Rate
3. First Serve Win Rate

## Clay Court

The most statistically significant variables for grass court were
1. Break Point Attempt Rate
2. Break Point Faced Rate
3. Second Serve Rate

## Exmple of Interpreting a coefficent

My example: Break points faced rate for the clay courts multivariable 
regression

Holding all other variables constant, someone would expect as the mean break 
point faced rate increases by one percent, the average winning percentage of a 
player would decrease by 3.81 percentage points.


# Errors

## Possible Reasons for High P-values in Some Variables


```{r}
#rcorr(as.matrix(atp_data_player[2:13]))
#rcorr(as.matrix(atp_data_hard[2:13]))
#rcorr(as.matrix(atp_data_grass[2:13]))
#rcorr(as.matrix(atp_data_clay[2:13]))

# Used to Create Correlation Coefficent Table

```


As you can see in the correlation table, there was a high correlation between
break points faced and second serve rate (-0.85 for hard, -0.69 for grass, 
and -0.73 for clay). This makes sense because one would expect that if someone 
wins a lot of their first serve points they would not face many break points 
and vice versa. This high correlation could cause their to be a high p-value 
in mean break points faced for hard courts, and first serve win rate for clay. 

As for the 0.28 p-value for second serve rate in grass court, there was not
really any other variables in the regression that it was highly associated with;
hence, this obscurity is some we would further investigate if given more time.

## Residual Plots

```{r}

# Only show residual vs fitted

plot(multi_lm_atp_clay,
     pch = 19, 
     col = "red",
     which = 1)

plot(multi_lm_atp_grass, 
     col = "green",
     pch = 19,
     which = 1)

plot(multi_lm_atp_hard, 
     col = "blue",
     pch = 19,
     which = 1)

```

In our residual plot (Fitted Values vs Residuals) most of the values were evenly
spread out, above and below zero, and the trend line is relatively straight,
showing that a linear model is probably appropriate. However, in each plot
there is a straight diagonal line of points. These are actually values of 
players who had a mean wining percentage of zero. Moreover, typically these 
players were lower ranked and as a result had to play the best players in world 
and hence, played fewer matches. We considered removing these players, but we 
decided not to because we had already limited our data to players in the top 100.


# Problems with our Analysis and What we Would do given More Time

The largest problem with our analysis is that we had no groundstroke, net
(i.e. approach shots, volley, and overhead), and other stroke data. Furthermore,
we did not have tracking data, and in tennis tracking data is extremely 
important, because footwork, being one of, if not, the most important aspects 
of tennis, can only be analyzed given tracking data. 

Some possible confounding variables in our study are the fact that most of data
are from grandslams or ATP Masters (1000) tournaments. Another possible 
confounding is dominant hand. We wanted to separate the data by hand, but we
realized that only 21 of the 128 players used in the analysis
were left handed, thus we could not make accurate interpretations of the data.

For further explorations, we will look at the obscurities in the findings
in this study, find a more complete data set, and adjust for more confounding
variables.


